---
title: "Analysis for Collins et al. Citizen Science Theory and Practice manuscript"
author: "Keith Jennings"
date: "`r Sys.Date()`"
output: github_document
---

Import the tidyverse package and the study data.

```{r}
library(tidyverse)
df <- read.csv("../../data/processed/mros_obs_gpm_processed_20220503.csv")
```

Filter to just the study domain and obs that pass QC.

```{r}
studyDomain = data.frame(ecoregions = c("High Plains", 
                                        "Southern Rockies", 
                                        "Colorado Plateaus",   # Colorado
                                        "Central Basin and Range", 
                                        "Sierra Nevada",               # Sierra
                                        "Willamette Valley", 
                                        "Cascades",                          # Oregon
                                        "Eastern Great Lakes Lowlands", 
                                        "Northeastern Highlands"), # Northeast
                         regions = c(rep("Colorado", 3),
                                   rep("Sierra", 2),
                                   rep("Oregon", 2),
                                   rep("Northeast", 2)))

obsStudy <- filter(df, 
                   eco_l3 %in% studyDomain$ecoregion)
obsPass <- filter(obsStudy, 
                  tair_flag == "Pass" &
                  ppt_flag == "Pass" & 
                  rh_flag == "Pass" &
                  dist_flag == "Pass" & 
                  dupe_flag == "Pass") %>% 
  left_join(., studyDomain,
            by = c("eco_l3" = "ecoregions"))
```

Some summary info:

- There were `r length(df$id)` observations during the study period from `r min(df$date)` to `r max(df$date)`
- Of those observations, `r length(obsStudy$id)` were in the targeted study regions
- `r length(obsPass$id)` observations (`r round(length(obsPass$id)/length(obsStudy$id) * 100, digits = 1)`%) in the study regions passed our quality control checks

Observations by Level 3 Ecoregion

```{r echo = FALSE, warning = FALSE, message = FALSE}
obsPass %>% 
  group_by(eco_l3) %>% 
  summarise(n = n()) %>% 
  arrange(desc(n)) %>% 
  knitr::kable()
```

Observations by study region

```{r echo = FALSE, warning = FALSE, message = FALSE}
obsPass %>% 
  group_by(regions) %>% 
  summarise(n = n()) %>% 
  arrange(desc(n)) %>% 
  knitr::kable()
```
