---
title: "Text and Analysis for in-progress Mountain Rain or Snow Manuscript"
author: "Keith Jennings"
date: "Sys.Date()"
output:
  github_document: default
---

# Info

The directory, for which this document serves as the README, includes all of the R scripts for tuning, processing, and analyzing rain-snow partitioning models.

<DETAIL THE FILES>

# The necessary stuff

Import the packages we need.

```{r message=FALSE}

# Load R packages
library(tidyverse)
library(tidymodels)
library(cowplot); theme_set(theme_cowplot())
```

Quantify the number of obs and date range for the crowdsourced data.

```{r}
cs_df <- read.csv("../../../../data/mros_QAQCflags_2024_01_24.csv")
cs_df <- cs_df %>% 
  filter(dupe_flag == "Pass",
         phase_flag == "Pass",
         temp_air_flag == "Pass",
         rh_flag == "Pass",
         nstation_temp_air_flag == "Pass",
         nstation_temp_dew_flag == "Pass",
  )

# Perform alternative dupe check to catch any that snuck through
cs_df  <- cs_df %>% 
  group_by(latitude, longitude, datetime_utc) %>% 
  mutate(n_obs = n()) %>% 
  filter(n_obs == 1)

# Format date and compute date range
cs_df$datetime <- ymd_hm(cs_df$datetime_utc)
range(cs_df$datetime, na.rm = T)
```


Import the datasets we need.

```{r message=FALSE}

# `cache=TRUE` keeps the data stored and prevents it from being read in anew each time

# Data directory prefix
# Specific to my computer, change before sharing
data_pre <- "../../../../data/"

# Load the results files
# These are local files (update document if moved)
# Also update nh scenario to be "nomix_onlymet"
# And add a source_label for plotting & tables
summary_all <- readRDS(paste0(data_pre, "performance_summary_all.RDS")) %>% 
    mutate(source_lab = ifelse(source == "cs", "Crowdsourced", "Synoptic"))
summary_bytemp <- readRDS(paste0(data_pre, "performance_summary_bytemp.RDS")) %>% 
    mutate(source_lab = ifelse(source == "cs", "Crowdsourced", "Synoptic"))
predictions_all <- readRDS(paste0(data_pre, "predictions_all.RDS"))

# Load the model objects for xgboost and random forest
ml_for_vip <- readRDS(paste0(data_pre, "ml_models_for_vip.RDS"))

# Load the met data for analysis
met <- readRDS(paste0(data_pre, "cs_nh_met_data.RDS"))

```

# Results

## Rain-snow partitioning performance

### Benchmarks

```{r}
benchmark_types <- data.frame(ppm = c("binlog","thresh_tdew_0", "thresh_tdew_0.5",
                                      "thresh_twet_0", "thresh_twet_0.5", 
                                      "thresh_twet_1",
                                      "thresh_tair_1", "thresh_tair_1.5",
                                      "rf", "xg", "nn"),
                              type = c(rep("humidity", 6), 
                                       rep("tair", 2),
                                       rep("ml", 3)))
summary_all <- left_join(summary_all, benchmark_types, by = "ppm")

```

The benchmark methods exhibited variable performance in predicting rain versus snow in both the crowdsourced and synoptic datasets (Table 1). The methods incorporating humidity (e.g., the binary logistic regression model and the wet bulb and dew point temperature thresholds) generally performed better than the methods using only air temperature, with a median accuracy value of `r filter(summary_all, type == "humidity") %>% pull(accuracy_pct) %>% median() %>% round(digits = 1)`% for the former compared to `r filter(summary_all, type == "tair") %>% pull(accuracy_pct) %>% median() %>% round(digits = 1)`% for the latter. With a median rain bias value of `r filter(summary_all, type == "humidity") %>% pull(rain_bias_pct) %>% median() %>% round(digits = 1)`%, the humidity-based methods performed markedly better than the air-temperature-only methods (`r filter(summary_all, type == "tair") %>% pull(rain_bias_pct) %>% median() %>% round(digits = 1)`%). The air-temperature-only methods also had a slightly worse median snow bias than the humidity methods at `r filter(summary_all, type == "tair") %>% pull(snow_bias_pct) %>% median() %>% round(digits = 1)`% versus `r filter(summary_all, type == "humidity") %>% pull(snow_bias_pct) %>% median() %>% round(digits = 1)`%.

```{r}
# Calculate minimums, maximums, and ranges

# Crowdsourced
# Accuracy
cs_benchmark_accuracy_min = summary_all %>% filter(source == "cs", scenario == "nomix_imbal" & !ppm %in% c("xg", "rf")) %>% pull(accuracy_pct) %>% min() %>% round(digits = 1)
cs_benchmark_accuracy_max = summary_all %>% filter(source == "cs", scenario == "nomix_imbal" & !ppm %in% c("xg", "rf")) %>% pull(accuracy_pct) %>% max() %>% round(digits = 1)
cs_benchmark_accuracy_range = cs_benchmark_accuracy_max - cs_benchmark_accuracy_min
# Snow bias
cs_benchmark_snow_bias_min = summary_all %>% filter(source == "cs", scenario == "nomix_imbal" & !ppm %in% c("xg", "rf")) %>% pull(snow_bias_pct) %>% min() %>% round(digits = 1)
cs_benchmark_snow_bias_max = summary_all %>% filter(source == "cs", scenario == "nomix_imbal" & !ppm %in% c("xg", "rf")) %>% pull(snow_bias_pct) %>% max() %>% round(digits = 1)
cs_benchmark_snow_bias_range = cs_benchmark_snow_bias_max - cs_benchmark_snow_bias_min
# Rain bias
cs_benchmark_rain_bias_min = summary_all %>% filter(source == "cs", scenario == "nomix_imbal" & !ppm %in% c("xg", "rf")) %>% pull(rain_bias_pct) %>% min() %>% round(digits = 1)
cs_benchmark_rain_bias_max = summary_all %>% filter(source == "cs", scenario == "nomix_imbal" & !ppm %in% c("xg", "rf")) %>% pull(rain_bias_pct) %>% max() %>% round(digits = 1)
cs_benchmark_rain_bias_range = cs_benchmark_rain_bias_max - cs_benchmark_rain_bias_min


# Synoptic
nh_benchmark_accuracy_min = summary_all %>% filter(source == "nh", scenario == "nomix_imbal" & !ppm %in% c("xg", "rf")) %>% pull(accuracy_pct) %>% min() %>% round(digits = 1)
nh_benchmark_accuracy_max = summary_all %>% filter(source == "nh", scenario == "nomix_imbal" & !ppm %in% c("xg", "rf")) %>% pull(accuracy_pct) %>% max() %>% round(digits = 1)
nh_benchmark_accuracy_range = nh_benchmark_accuracy_max - nh_benchmark_accuracy_min
# Snow bias
nh_benchmark_snow_bias_min = summary_all %>% filter(source == "nh", scenario == "nomix_imbal" & !ppm %in% c("xg", "rf")) %>% pull(snow_bias_pct) %>% min() %>% round(digits = 1)
nh_benchmark_snow_bias_max = summary_all %>% filter(source == "nh", scenario == "nomix_imbal" & !ppm %in% c("xg", "rf")) %>% pull(snow_bias_pct) %>% max() %>% round(digits = 1)
nh_benchmark_snow_bias_range = nh_benchmark_snow_bias_max - nh_benchmark_snow_bias_min
# Rain bias
nh_benchmark_rain_bias_min = summary_all %>% filter(source == "nh", scenario == "nomix_imbal" & !ppm %in% c("xg", "rf")) %>% pull(rain_bias_pct) %>% min() %>% round(digits = 1)
nh_benchmark_rain_bias_max = summary_all %>% filter(source == "nh", scenario == "nomix_imbal" & !ppm %in% c("xg", "rf")) %>% pull(rain_bias_pct) %>% max() %>% round(digits = 1)
nh_benchmark_rain_bias_range = nh_benchmark_rain_bias_max - nh_benchmark_rain_bias_min


```

When examining the two datasets separately, we found a few key differences. For example, accuracy was generally lower in the crowdsourced dataset, ranging from `r cs_benchmark_accuracy_min`% to `r cs_benchmark_accuracy_max`%, a total spread of `r cs_benchmark_accuracy_range`%. That is larger than the accuracy spread of the synoptic dataset (`r nh_benchmark_accuracy_range`%), which had a minimum of `r nh_benchmark_accuracy_min`% and a maximum of `r nh_benchmark_accuracy_max`%. In the crowdsourced dataset, the top six methods, ranked by accuracy, all used humidity, while the bottom two relied on air temperature alone. The case was slightly different for the synoptic dataset where the two dew point temperature thresholds performed worse than the two static air temperature thresholds. Snow and rain biases had larger ranges than accuracy in both datasets. Snow bias varied by `r cs_benchmark_snow_bias_range`% and `r nh_benchmark_snow_bias_range`%, while rain bias varied by `r cs_benchmark_rain_bias_range`% and `r nh_benchmark_rain_bias_range`% in the crowdsourced and synoptic datasets, respectively. In general, the benchmarks performed better and with less variability when applied to the synoptic dataset, and methods with better accuracy values typically had lower rain and snow biases.

```{r}
summary_all %>% 
    filter(scenario == "nomix_imbal") %>% 
    filter(type != "ml") %>% 
    ungroup() %>% 
    arrange(source_lab, -accuracy_pct) %>% 
    mutate(across(where(is.numeric), round, digits = 1)) %>% 
    select(Source = source_lab, 
           PPM = ppm, 
           `Accuracy (%)` = accuracy_pct, 
           `Snow bias (%)` = snow_bias_pct,
           `Rain bias (%)` = rain_bias_pct) %>% 
    knitr::kable()
```

Table: Table 1. Performance metrics for the benchmark precipitation partitioning methods (PPMs) using the crowdsourced and synoptic datasets.

All of the benchmark methods performed worse at air temperatures near and slightly above the freezing point in terms of accuracy, with most of them reaching their minimum accuracy values between 1.5°C and 2.0°C in the crowdsourced dataset and between 0.5°C and 1.5°C in the synoptic dataset (Figure 3). The best performing benchmarks in terms of accuracy were T~w1.0~ and T~w0.5~ in the crowdsourced and synoptic datasets, respectively. For the former, T~w1.0~ hit a minimum accuracy of `r summary_bytemp %>% filter(scenario == "nomix_imbal" & source == "cs" & ppm == "thresh_twet_1") %>% slice_min(accuracy_pct) %>% pull(accuracy_pct) %>% round(digits = 1)`% at `r summary_bytemp %>% filter(scenario == "nomix_imbal" & source == "cs" & ppm == "thresh_twet_1") %>% slice_min(accuracy_pct) %>% pull(tair_bin) %>% round(digits = 1)`°C, while the latter reached a minimum accuracy of `r summary_bytemp %>% filter(scenario == "nomix_imbal" & source == "nh" & ppm == "thresh_twet_0.5") %>% slice_min(accuracy_pct) %>% pull(accuracy_pct) %>% round(digits = 1)`% at `r summary_bytemp %>% filter(scenario == "nomix_imbal" & source == "nh" & ppm == "thresh_twet_0.5") %>% slice_min(accuracy_pct) %>% pull(tair_bin) %>% round(digits = 1)`°C. In both datasets, most methods had slightly positive snow biases and markedly negative rain biases at air temperatures less than 1.0°C. This is consistent with their limited, or complete lack of, ability to predict rain at sub-freezing temperatures. Conversely, most methods, with the exception of the dew point temperature thresholds, had largely negative snow biases and slightly positive rain biases at air temperatures above 2°C. Notably, the crowdsourced dataset expressed larger positive rain biases at higher temperatures, indicating the benchmark methods failed to capture snowfall occurring during warm near-surface conditions.

```{r benchmark_plot, echo=F, message=F, fig.cap = "", fig.height=4, fig.width=6}

# Make PPM labels
ppm_labels <- 
    data.frame(
        ppm = c(
            "binlog",
            "nn",
            "rf",
            "thresh_tair_1",
            "thresh_tair_1.5",
            "thresh_tdew_0",
            "thresh_tdew_0.5",
            "thresh_twet_0",
            "thresh_twet_0.5",
            "thresh_twet_1",
            "xg"
        ),
        ppm_labs = c(
            "Bin[log]",
            "ANN",
            "RF",
            "T[a1.0]",
            "T[a1.5]",
            "T[d0.0]",
            "T[d0.5]",
            "T[w0.0]",
            "T[w0.5]",
            "T[w1.0]",
            "XG"
        )
    )

# Filter data to plot and join labels
plot_data <- summary_bytemp %>% 
    filter(scenario == "nomix_imbal") %>% 
    filter(!ppm %in% c("xg", "rf", "nn")) %>% 
    left_join(ppm_labels, by = "ppm") %>% 
    mutate(source2 = ifelse(source == "cs", "Crowdsourced", "Synoptic"))

# Make the plot
benchmark_accuracy_bias_plot <-
  plot_grid(
  ggplot(plot_data, aes(tair_bin, ppm_labs, fill = accuracy_pct)) + 
    geom_raster() + 
    geom_vline(xintercept = 0, lty = "dashed") +
    facet_wrap(~source2) + 
    xlim(-8, 8) +
    scale_y_discrete("PPM", labels = rev(parse(text = unique(plot_data$ppm_labs))),
                     limits = rev) +
    labs(x = "Air temperature (°C)") +
    #scale_fill_gradient(low="honeydew1", high="#009194", name = "Frequency (%)") +
    #scale_fill_viridis_c(option = "inferno", direction = -1) +
    scale_fill_gradientn(colors = hcl.colors(3, "Heat2")) +
    guides(fill = guide_colorbar(title.position = "top", title = "Accuracy (%)",
                                 barwidth = 10)) +
    theme(legend.position = "top", legend.justification = "center", 
          legend.title.align = 0.5) ,
  ggplot(plot_data, aes(tair_bin, ppm_labs, fill = snow_bias_pct)) + 
    geom_raster() + 
    geom_vline(xintercept = 0, lty = "dashed") +
    facet_wrap(~source2) + 
    xlim(-8, 8) +
    scale_fill_gradientn(limits = c(-100,100), oob = squish, na.value = "white", 
                         colors = hcl.colors(3, "Blue-Red2", rev = T))  +
    labs(x = "Air temperature (°C)")+
    guides(fill = guide_colorbar(title.position = "top", title = "Snow Bias (%)",
                                 barwidth = 10)) +
    theme(legend.position = "top", legend.justification = "center",
          legend.title.align = 0.5,
          axis.title.y = element_blank(),
          axis.text.y = element_blank()),
  ggplot(plot_data, aes(tair_bin, ppm_labs, fill = rain_bias_pct)) + 
    geom_raster() + 
    geom_vline(xintercept = 0, lty = "dashed") +
    facet_wrap(~source2) + 
    xlim(-8, 8) +
    scale_fill_gradientn(limits = c(-100,100), oob = squish, na.value = "white", 
                         colors = hcl.colors(3, "Blue-Red2", rev = T)) +
    labs(x = "Air temperature (°C)") +
    guides(fill = guide_colorbar(title.position = "top", title = "Rain Bias (%)",
                                 barwidth = 10)) +
    theme(legend.position = "top", legend.justification = "center",
          legend.title.align = 0.5,
          axis.title.y = element_blank(),
          axis.text.y = element_blank()),
  nrow = 1, rel_widths = c(1, 0.8, 0.8), labels = "auto"
)

benchmark_accuracy_bias_plot
save_plot(filename = "figures/fig03_benchmark_accuracy_bias.png", 
          plot = benchmark_accuracy_bias_plot, 
          base_height = 5, 
          base_width = 9)
```

```{r}
# Code for minimum accuracy
summary_bytemp %>% filter(source == "cs" & scenario == "nomix_imbal" & !ppm %in% c("rf", "xg", "nn")) %>% group_by(ppm) %>% slice_min(accuracy_pct)
summary_bytemp %>% filter(source == "nh" & scenario == "nomix_imbal" & !ppm %in% c("rf", "xg", "nn")) %>% group_by(ppm) %>% slice_min(accuracy_pct)

```


### Machine learning methods

```{r}
# Identify columns to select
cols_to_get = c("source", "ppm", "accuracy_pct", "snow_bias_pct", "rain_bias_pct")

# Get the best benchmark in crowdsourced data
cs_best_benchmark = summary_all %>% 
    filter(source == "cs" & !ppm %in% c("xg", "rf", "nn") & scenario == "nomix_imbal") %>% 
    ungroup() %>% 
    slice_max(accuracy_pct, n = 1) %>% 
    pull(ppm)

# Get the average accuracy 
cs_accuracy_av <- summary_all %>% 
  filter(scenario == "nomix_imbal" & source == "cs" & !ppm %in% c("rf", "xg", "nn")) %>% 
  ungroup() %>% 
  summarize(accuracy_pct = mean(accuracy_pct),
            rain_bias_pct = mean(rain_bias_pct),
            snow_bias_pct = mean(snow_bias_pct)) %>% 
  mutate(source = "cs", "ppm" = "avg") %>% 
  select(cols_to_get)

# Extract and summarize results
cs_benchmark_diff <- summary_all %>% ungroup() %>% 
    filter(source == "cs" & ppm %in% c("xg", "rf", "nn") & scenario == "nomix_imbal") %>% 
    select(cols_to_get) %>% 
    bind_cols(.,summary_all %>% ungroup() %>% 
                  filter(source == "cs" & ppm == cs_best_benchmark & scenario == "nomix_imbal") %>% 
                  select(cols_to_get)) %>% 
    bind_rows(., summary_all %>% ungroup() %>% 
    filter(source == "cs" & ppm %in% c("xg", "rf", "nn") & scenario == "nomix_imbal") %>% 
    select(cols_to_get) %>% 
    bind_cols(., cs_accuracy_av))

# Get the best benchmark in synoptic data
nh_best_benchmark = summary_all %>% 
    filter(source == "nh" & !ppm %in% c("xg", "rf", "nn") & scenario == "nomix_imbal") %>% 
    ungroup() %>% 
    slice_max(accuracy_pct, n = 1) %>% 
    pull(ppm)

# Get the average accuracy 
nh_accuracy_av <- summary_all %>% 
  filter(scenario == "nomix_imbal" & source == "nh" & !ppm %in% c("rf", "xg", "nn")) %>% 
  ungroup() %>% 
  summarize(accuracy_pct = mean(accuracy_pct),
            rain_bias_pct = mean(rain_bias_pct),
            snow_bias_pct = mean(snow_bias_pct)) %>% 
  mutate(source = "cs", "ppm" = "avg") %>% 
  select(cols_to_get)

# Extract and summarize results
nh_benchmark_diff <- summary_all %>% ungroup() %>% 
    filter(source == "nh" & ppm %in% c("xg", "rf", "nn") & scenario == "nomix_imbal") %>% 
    select(cols_to_get) %>% 
    bind_cols(.,summary_all %>% ungroup() %>% 
                  filter(source == "nh" & ppm == nh_best_benchmark & scenario == "nomix_imbal") %>% 
                  select(cols_to_get)) %>% 
    bind_rows(., summary_all %>% ungroup() %>% 
    filter(source == "nh" & ppm %in% c("xg", "rf", "nn") & scenario == "nomix_imbal") %>% 
    select(cols_to_get) %>% 
    bind_cols(., nh_accuracy_av))

ml_benchmark_diff <- bind_rows(cs_benchmark_diff, nh_benchmark_diff) %>% 
    mutate(comparison = rep(c("best", "best", "best", "avg", "avg", "avg"), 2)) %>% 
    select(source = source...1,
           ppm = ppm...2,
           comparison,
           ml_accuracy_pct = accuracy_pct...3,
           ml_snow_bias_pct = snow_bias_pct...4,
           ml_rain_bias_pct = rain_bias_pct...5,
           bm_accuracy_pct = accuracy_pct...8,
           bm_snow_bias_pct = snow_bias_pct...9,
           bm_rain_bias_pct = rain_bias_pct...10) %>% 
    mutate(accuracy_pct_diff = ml_accuracy_pct - bm_accuracy_pct,
           snow_bias_abs_pct_diff = abs(ml_snow_bias_pct) - abs(bm_snow_bias_pct),
           rain_bias_abs_pct_diff = abs(ml_rain_bias_pct) - abs(bm_rain_bias_pct))

# Make a table
ml_benchmark_diff %>% 
    select(source, ppm, comparison,
           ml_accuracy_pct, ml_snow_bias_pct, ml_rain_bias_pct, 
           accuracy_pct_diff, snow_bias_abs_pct_diff, rain_bias_abs_pct_diff) %>% 
    mutate(across(where(is.numeric), round, digits = 1)) %>% 
    knitr::kable()

``` 

The three machine learning methods, ANN, random forest and XGBoost, generally provided small performance gains relative to the best benchmark methods and larger, more consistent improvements relative to the average benchmark results (Table 2). In the crowdsourced dataset, ANN (`r summary_all %>% filter(source == "cs" & ppm == "nn" & scenario == "nomix_imbal") %>% pull(accuracy_pct) %>% round(digits = 1)`%) and XGBoost (`r summary_all %>% filter(source == "cs" & ppm == "xg" & scenario == "nomix_imbal") %>% pull(accuracy_pct) %>% round(digits = 1)`%) had slightly higher accuracy values than T~w1.0~ (`r summary_all %>% filter(source == "cs" & ppm == "thresh_twet_1" & scenario == "nomix_imbal") %>% pull(accuracy_pct) %>% round(digits = 1)`%), while random forest had a slightly lower accuracy (`r summary_all %>% filter(source == "cs" & ppm == "rf" & scenario == "nomix_imbal") %>% pull(accuracy_pct) %>% round(digits = 1)`%) than the T~w1.0~ threshold, which was the best performing method from the benchmark exercise. Both XGBoost (`r summary_all %>% filter(source == "nh" & ppm == "xg" & scenario == "nomix_imbal") %>% pull(accuracy_pct) %>% round(digits = 1)`%) and random forest (`r summary_all %>% filter(source == "nh" & ppm == "rf" & scenario == "nomix_imbal") %>% pull(accuracy_pct) %>% round(digits = 1)`%) provided marginal improvements relative to T~w0.5~ (`r summary_all %>% filter(source == "nh" & ppm == "thresh_twet_0.5" & scenario == "nomix_imbal") %>% pull(accuracy_pct) %>% round(digits = 1)`%), the best benchmark in the synoptic dataset as measured by the accuracy metric. ANN, however, expressed a slightly worse accuracy value (`r summary_all %>% filter(source == "nh" & ppm == "nn" & scenario == "nomix_imbal") %>% pull(accuracy_pct) %>% round(digits = 1)`%) than the best benchmark.

```{r}
# Code for minimum accuracy
summary_bytemp %>% filter(source == "cs" & scenario == "nomix_imbal" & ppm %in% c("rf", "xg", "nn")) %>% group_by(ppm) %>% slice_min(accuracy_pct)
```


```{r}
# Code for minimum accuracy
summary_bytemp %>% filter(source == "nh" & scenario == "nomix_imbal" & ppm %in% c("rf", "xg", "nn")) %>% group_by(ppm) %>% slice_min(accuracy_pct)
```

```{r}
# Add average values to summary by temp
performance_av_bytemp <- summary_bytemp %>% 
  filter(scenario == "nomix_imbal" & !ppm %in% c("rf", "xg", "nn")) %>% 
  group_by(tair_bin, source) %>% 
  summarize(accuracy_pct = mean(accuracy_pct),
            rain_bias_pct = mean(rain_bias_pct),
            snow_bias_pct = mean(snow_bias_pct)) %>% 
  mutate(ppm = "avg", scenario = "nomix_imbal")

# Bind to summary_bytemp
summary_bytemp_withavg <- bind_rows(summary_bytemp,
                                    performance_av_bytemp)

# Summarize for each dataset
ml_benchmark_diff_bytemp <- bind_rows(
  summary_bytemp_withavg %>% 
    filter(source == "cs" & scenario == "nomix_imbal") %>% 
    pivot_wider(names_from = ppm, values_from = accuracy_pct, id_cols = tair_bin) %>% 
    mutate(rf_best = (rf-thresh_twet_1)/thresh_twet_1*100,
           rf_avg = (rf-avg)/avg*100,
           xg_best = (xg-thresh_twet_1)/thresh_twet_1*100,
           xg_avg = (xg-avg)/avg*100,
           nn_best = (nn-thresh_twet_1)/thresh_twet_1*100,
           nn_avg = (nn-avg)/avg*100) %>% 
    select(tair_bin, rf_best:nn_avg) %>% 
    pivot_longer(cols = rf_best:nn_avg,
                 names_to = c("ppm", "benchmark"),
                 names_pattern = "(.*)_(.*)",
                 values_to = "accuracy_diff_rel_pct") %>% 
    mutate(source_lab = "Crowdsourced", source = "cs"),
  summary_bytemp_withavg %>% 
    filter(source == "nh" & scenario == "nomix_imbal") %>% 
    pivot_wider(names_from = ppm, values_from = accuracy_pct, id_cols = tair_bin) %>% 
    mutate(rf_best = (rf-thresh_twet_0.5)/thresh_twet_0.5*100,
           rf_avg = (rf-avg)/avg*100,
           xg_best = (xg-thresh_twet_0.5)/thresh_twet_0.5*100,
           xg_avg = (xg-avg)/avg*100,
           nn_best = (nn-thresh_twet_0.5)/thresh_twet_0.5*100,
           nn_avg = (nn-avg)/avg*100) %>% 
    select(tair_bin, rf_best:nn_avg) %>% 
    pivot_longer(cols = rf_best:nn_avg,
                 names_to = c("ppm", "benchmark"),
                 names_pattern = "(.*)_(.*)",
                 values_to = "accuracy_diff_rel_pct") %>% 
    mutate(source_lab = "Synoptic", source = "nh")
)

ml_benchmark_rel_diff_plot <- 
  ggplot() +
  geom_hline(yintercept = 0, color = "grey") +
  geom_line(data = ml_benchmark_diff_bytemp, lwd = 1, 
            aes(tair_bin, accuracy_diff_rel_pct, color = ppm, lty = benchmark)) +
  scale_color_manual(values =  c("#4a6fe3", "gray26", "#d33f6a"),
                     labels = c("ANN","RF", "XG"),
                     name = "PPM")+
  scale_linetype_discrete(labels = c("Average", "Best"),
                          name = "Benchmark\ncomparison") +
  labs(x = "Air temperature (°C)", 
       y = "Relative accuracy difference to benchmark (%)") +
  xlim(c(-5,10)) +
  theme(legend.position = c(0.03,0.7)) +
  facet_wrap(~source_lab)
ml_benchmark_rel_diff_plot
save_plot(filename = "figures/fig02_ml_benchmark_rel_diff.png", 
          plot = ml_benchmark_rel_diff_plot, 
          base_height = 5, 
          base_width = 9)


```

The three machine learning methods exhibited similar accuracy patterns by air temperature as the best benchmark and the average benchmark results, with performance degrading at air temperatures slightly above freezing. Both random forest and XGBoost achieved their minimum success rates of `r filter(summary_bytemp, source == "cs" & scenario == "nomix_imbal" & ppm == "rf") %>% pull(accuracy_pct) %>% min() %>% round(digits = 1)`% and `r filter(summary_bytemp, source == "cs" & scenario == "nomix_imbal" & ppm == "xg") %>% pull(accuracy_pct) %>% min() %>% round(digits = 1)`%, respectively, at 2°C in the crowdsourced dataset. Meanwhile, ANN had its worst accuracy value of `r filter(summary_bytemp, source == "cs" & scenario == "nomix_imbal" & ppm == "nn") %>% pull(accuracy_pct) %>% min() %>% round(digits = 1)`% at 2.5°C. Performance was higher in the synoptic dataset for random forest and XGBoost with minimum accuracy values of `r filter(summary_bytemp, source == "nh" & scenario == "nomix_imbal" & ppm == "rf") %>% pull(accuracy_pct) %>% min() %>% round(digits = 1)`% and `r filter(summary_bytemp, source == "nh" & scenario == "nomix_imbal" & ppm == "xg") %>% pull(accuracy_pct) %>% min() %>% round(digits = 1)`%, respectively. ANN, conversely, expressed slightly worse performance than in the crowdsourced dataset with a minimum accuracy of `r filter(summary_bytemp, source == "nh" & scenario == "nomix_imbal" & ppm == "nn") %>% pull(accuracy_pct) %>% min() %>% round(digits = 1)`%. All minima occurred at 1°C in the synoptic dataset. Relative improvements in accuracy compared to the best and average benchmarks varied by temperature for the machine learning methods (Figure 2). In the crowdsourced dataset, relative improvements reached a maximum of `r filter(ml_benchmark_diff_bytemp, source == "cs" & benchmark == "best") %>% pull(accuracy_diff_rel_pct) %>% max() %>% round(digits = 1)`% at `r filter(ml_benchmark_diff_bytemp, source == "cs" & benchmark == "best") %>% slice_max(accuracy_diff_rel_pct) %>% pull(tair_bin) %>% round(digits = 1) `°C compared to the best benchmark and a maximum of `r filter(ml_benchmark_diff_bytemp, source == "cs" & benchmark == "avg") %>% pull(accuracy_diff_rel_pct) %>% max() %>% round(digits = 1)`% at `r filter(ml_benchmark_diff_bytemp, source == "cs" & benchmark == "avg") %>% slice_max(accuracy_diff_rel_pct) %>% pull(tair_bin) %>% round(digits = 1) `°C compared to the average benchmark values. In the synoptic dataset relative improvements were similar for the best benchmark comparison, with a max increase of `r filter(ml_benchmark_diff_bytemp, source == "nh" & benchmark == "best") %>% pull(accuracy_diff_rel_pct) %>% max() %>% round(digits = 1)`% at `r filter(ml_benchmark_diff_bytemp, source == "nh" & benchmark == "best") %>% slice_max(accuracy_diff_rel_pct) %>% pull(tair_bin) %>% round(digits = 1) `°C, and smaller for the average benchmark values, with a max increase of `r filter(ml_benchmark_diff_bytemp, source == "nh" & benchmark == "avg") %>% pull(accuracy_diff_rel_pct) %>% max() %>% round(digits = 1)`% at `r filter(ml_benchmark_diff_bytemp, source == "nh" & benchmark == "avg") %>% slice_max(accuracy_diff_rel_pct) %>% pull(tair_bin) %>% round(digits = 1) `°C. Notably, ANN did not provide an accuracy improvement at any air temperature relative to the best benchmark in the synoptic dataset. 

When compared to the snow and rain biases produced by the best benchmarks and average benchmark values, the machine learning methods generally provided larger absolute improvements than those recorded for accuracy (Table 2). While accuracy improvements were all less than or equal to `r max(ml_benchmark_diff$accuracy_pct_diff) %>% round(digits = 1)`%, the reduction in snow bias maxed out at `r filter(ml_benchmark_diff, source == "cs") %>% pull(snow_bias_abs_pct_diff) %>% min() %>% round(digits = 1)`% and `r filter(ml_benchmark_diff, source == "nh") %>% pull(snow_bias_abs_pct_diff) %>% min() %>% round(digits = 1)`% in the crowdsourced and synoptic datasets, respectively. Random forest, XGBoost, and ANN also provided consistent improvements in rain bias, with a maximum reduction of `r filter(ml_benchmark_diff, source == "cs") %>% pull(rain_bias_abs_pct_diff) %>% min() %>% round(digits = 1)`% in the crowdsourced dataset and `r filter(ml_benchmark_diff, source == "nh") %>% pull(rain_bias_abs_pct_diff) %>% min() %>% round(digits = 1)`% in the synoptic dataset. These results show that while accuracy improvements from the machine learning methods may be small, random forest, XGBoost, and ANN predict rain and snow with lower biases than the benchmark methods.

## Mixed precipitation

Until this point, we have only considered precipitation in its solid and liquid forms. While the synoptic dataset does not include any mixed precipitation observations, the testing split of the crowdsourced dataset includes `r filter(met, source == "cs" & scenario == "allphase_imbal" & phase == "mix") %>% nrow()` mixed phase observations, comprising `r ((filter(met, source == "cs" & scenario == "allphase_imbal" & phase == "mix") %>% nrow()) / (filter(met, source == "cs" & scenario == "allphase_imbal") %>% nrow()) * 100) %>% round(digits = 1)`% of the total. As noted in the methods section, we do not include any dual-threshold benchmark methods because of their poor historical performance. The machine learning methods analyzed here fared little better. Including the mixed observations caused overall accuracy to markedly decline, going from `r summary_all %>% filter(source == "cs" & ppm == "rf" & scenario == "nomix_imbal") %>% pull(accuracy_pct) %>% round(digits = 1)`% (rain and snow only) to `r summary_all %>% filter(source == "cs" & ppm == "rf" & scenario == "allphase_imbal") %>% pull(accuracy_pct) %>% round(digits = 1)`% (rain, snow, and mixed) for random forest, from `r summary_all %>% filter(source == "cs" & ppm == "xg" & scenario == "nomix_imbal") %>% pull(accuracy_pct) %>% round(digits = 1)`% (rain and snow only) to `r summary_all %>% filter(source == "cs" & ppm == "xg" & scenario == "allphase_imbal") %>% pull(accuracy_pct) %>% round(digits = 1)`% (rain, snow, and mixed) for XGBoost, and from `r summary_all %>% filter(source == "cs" & ppm == "nn" & scenario == "nomix_imbal") %>% pull(accuracy_pct) %>% round(digits = 1)`% (rain and snow only) to `r summary_all %>% filter(source == "cs" & ppm == "nn" & scenario == "allphase_imbal") %>% pull(accuracy_pct) %>% round(digits = 1)`% (rain, snow, and mixed) for ANN. 

Although XGBoost and ANN had higher accuracy values than random forest when we included mixed precipitation, they achieved this result at the expense of bias. The random forest method predicted that `r summary_all %>% filter(source == "cs" & ppm == "rf" & scenario == "allphase_imbal") %>% pull(mixed_pred_pct) %>% round(digits = 1)`% of the observations were mixed, which was well short of the observed value of `r summary_all %>% filter(source == "cs" & ppm == "rf" & scenario == "allphase_imbal") %>% pull(mixed_obs_pct) %>% round(digits = 1)`%, giving it a mixed bias of `r summary_all %>% filter(source == "cs" & ppm == "rf" & scenario == "allphase_imbal") %>% pull(mixed_bias_pct) %>% round(digits = 1)`%. XGBoost and ANN fared even worse, predicting `r summary_all %>% filter(source == "cs" & ppm == "xg" & scenario == "allphase_imbal") %>% pull(mixed_pred_pct) %>% round(digits = 1)`% of the observations to be mixed, giving them both a mixed bias of `r summary_all %>% filter(source == "cs" & ppm == "xg" & scenario == "allphase_imbal") %>% pull(mixed_bias_pct) %>% round(digits = 1)`%. When including mixed precipitation, random forest had a lower snow bias (`r summary_all %>% filter(source == "cs" & ppm == "rf" & scenario == "allphase_imbal") %>% pull(snow_bias_pct) %>% round(digits = 1)`%) than XGBoost (`r summary_all %>% filter(source == "cs" & ppm == "xg" & scenario == "allphase_imbal") %>% pull(snow_bias_pct) %>% round(digits = 1)`%) and ANN (`r summary_all %>% filter(source == "cs" & ppm == "nn" & scenario == "allphase_imbal") %>% pull(snow_bias_pct) %>% round(digits = 1)`%). It was the same story for rain bias with XGBoost (`r summary_all %>% filter(source == "cs" & ppm == "xg" & scenario == "allphase_imbal") %>% pull(rain_bias_pct) %>% round(digits = 1)`%) and ANN  (`r summary_all %>% filter(source == "cs" & ppm == "nn" & scenario == "allphase_imbal") %>% pull(rain_bias_pct) %>% round(digits = 1)`%) having worse biases than random forest (`r summary_all %>% filter(source == "cs" & ppm == "rf" & scenario == "allphase_imbal") %>% pull(rain_bias_pct) %>% round(digits = 1)`%). These findings appear even worse when viewing the confusion matrix in Figure 3. This plot shows that random forest correctly predicted only 9.3% of the observed mixed precipitation and incorrectly predicted the mixed observations as rain 32.9% of the time and as snow the remaining 57.8%. XGBoost and ANN failed completely here, correctly predicting none of the mixed precipitation observations, identifying them as snow over 60% of the time. 

```{r}
# Get the mixed predictions
mixed_preds <- predictions_all %>%
  filter(scenario == "allphase_imbal" & ppm %in% c("rf", "xg", "nn")) %>% 
  mutate(phase_factor = as.factor(phase),
         phase_pred_factor = as.factor(phase_pred))

# Make a confusion matrix
mixed_conf_matrix_plot <- 
  bind_rows(
    mixed_preds %>% filter(ppm == "xg") %>%
      select(phase_pred_factor, phase_factor) %>%
      table() %>%
      prop.table(2) %>%
      as.data.frame() %>%
      mutate(ppm_lab = "XGBoost"),
    mixed_preds %>% filter(ppm == "rf") %>%
      select(phase_pred_factor, phase_factor) %>%
      table() %>%
      prop.table(2) %>%
      as.data.frame() %>%
      mutate(ppm_lab = "Random forest"),
     mixed_preds %>% filter(ppm == "nn") %>%
      select(phase_pred_factor, phase_factor) %>%
      table() %>%
      prop.table(2) %>%
      as.data.frame() %>%
      mutate(ppm_lab = "ANN")
  ) %>%
  ggplot(aes(phase_factor, phase_pred_factor, fill = Freq*100)) +
  geom_tile() +
  geom_text(aes(label = round(Freq * 100, digits = 1))) +
    scale_fill_gradientn(colors = hcl.colors(3, "Heat2"),
                         name = "Frequency (%)") +
  #scale_fill_gradient(low="honeydew1", high="#009194", name = "Frequency\n(%)") +
  labs(x = "Observed", y = "Predicted") +
  scale_x_discrete(labels = c("Mixed", "Rain", "Snow")) +
  scale_y_discrete(labels = c("Mixed", "Rain", "Snow")) +
  facet_wrap(~ppm_lab) +
  theme(legend.position = "top", legend.justification = "center",
          legend.title.align = 0.5) +
  guides(fill = guide_colorbar(title.position = "top", title = "Frequency (%)",
                                 barwidth = 10))

mixed_conf_matrix_plot

save_plot(filename = "figures/fig03_mixed_conf_matrix.png", 
          plot = mixed_conf_matrix_plot, 
          base_height = 5, 
          base_width = 9)

```


## Meteorology and snow frequency analysis

```{r}
frequency_data <- 
  bind_rows(
    summary_bytemp %>% 
      filter(scenario == "nomix_imbal" & tair_bin < 15 & tair_bin > -10) %>% 
      group_by(tair_bin, source) %>% 
      summarize(snow_pred_pct_min = min(snow_pred_pct),
                snow_pred_pct_max = max(snow_pred_pct),
                snow_pred_pct_avg = mean(snow_pred_pct), 
                snow_obs_pct = mean(snow_obs_pct),
                rain_pred_pct_min = min(rain_pred_pct),
                rain_pred_pct_max = max(rain_pred_pct),
                rain_pred_pct_avg = mean(rain_pred_pct), 
                rain_obs_pct = mean(rain_obs_pct)) %>% 
      mutate(snow_pred_diff = snow_pred_pct_max - snow_pred_pct_min,
             rain_pred_diff = rain_pred_pct_max - rain_pred_pct_min),
    summary_bytemp %>% 
      filter(scenario == "nomix_imbal" & tair_bin < 15 & tair_bin > -10) %>% 
      filter(ppm == "nn" & source == "cs" | ppm == "thresh_twet_1" & source == "cs" |
               ppm == "rf" & source == "nh" | ppm == "thresh_twet_0.5" & source == "nh") %>% 
      select(ppm, source, snow_pred_pct, rain_pred_pct, tair_bin) %>% 
      mutate(type = ifelse(source == "cs",
                           ifelse(ppm == "nn",
                                  "Machine learning",
                                  "Benchmark"),
                           ifelse(ppm == "rf",
                                  "Machine learning",
                                  "Benchmark")))
  ) %>% 
  mutate(source_lab = ifelse(source == "cs",
                             "Crowdsourced",
                             "Synoptic"),
         dummy = "dummy")

snowfall_freq_plot <- 
  ggplot() +
  geom_ribbon(data = filter(frequency_data, !is.na(snow_pred_pct_min) & !is.na(snow_pred_pct_max)),
              aes(tair_bin, ymin = snow_pred_pct_min, ymax = snow_pred_pct_max, fill = dummy), alpha = 0.7) +
  geom_line(data = filter(frequency_data, !is.na(snow_pred_pct)), 
            aes(tair_bin, snow_pred_pct, lty = type)) +
  geom_line(data = filter(frequency_data, !is.na(snow_obs_pct)), 
            aes(tair_bin, snow_obs_pct, color = dummy)) +
  scale_linetype_manual(values = c("dotted", "dashed"), name = "",
                        labels = c("Best benchmark", "Best ML")) + 
  scale_fill_manual(name = element_blank(), labels = "PPM range", values = 
                      "#d33f6a") +
  scale_color_manual(name = element_blank(), labels = "Observed", values = "black") +
  theme(legend.position = c(0.4, 0.5),
        legend.spacing.y = unit(-0.05, "cm")) +
  labs(x = "Air temperature (°C)", y = "Snowfall frequency (%)") + 
  facet_wrap(~source_lab, scales = "free_x")
  
snowfall_freq_plot

save_plot(filename = "figures/fig04_snowfall_freq.png", 
          plot = snowfall_freq_plot, 
          base_height = 4, 
          base_width = 9)


subzero_freq <- met %>% 
  filter(tair <= 0 & scenario == "nomix_imbal") %>% 
  group_by(source) %>% 
  summarize(n_rain = sum(phase == "rain"),
            n_snow = sum(phase == "snow"),
            n_all = n()) %>% 
  mutate(rain_pct = round(n_rain/n_all * 100, digits = 1))
```

In this section, we return our focus to the predictions of only rain and snow given the poor mixed phase results. Our first consideration is whether the benchmark and machine learning methods match observed snowfall frequency curves. A couple patterns emerge from Figure 6. One, all phase partitioning methods fail to capture the sub-freezing rainfall events observed in both datasets. Although rainfall comprises a small proportion of total precipitation below 0°C in the crowdsourced (`r filter(subzero_freq, source == "cs") %>% pull(rain_pct)`%) and synoptic (`r filter(subzero_freq, source == "nh") %>% pull(rain_pct)`%) datasets, these observations explain the rain biases approaching -100% shown in Figure 2 to the left of the 0°C isotherm. Two, the spreads between the minimum and maximum snowfall frequencies predicted by all of the phase partitioning methods (blue shading in Figure 6) approach their largest values at air temperatures slightly above 0°C. The largest difference (`r frequency_data %>% ungroup() %>% filter(source == "cs") %>% slice_max(snow_pred_diff) %>% pull(snow_pred_diff) %>% round(digits = 1)`%) occurs at `r frequency_data %>% ungroup() %>% filter(source == "cs") %>% slice_max(snow_pred_diff) %>% pull(tair_bin) %>% round(digits = 1)`°C in the crowdsourced dataset, which is the same temperature at which the synoptic dataset expresses its largest difference (`r frequency_data %>% ungroup() %>% filter(source == "nh") %>% slice_max(snow_pred_diff) %>% pull(snow_pred_diff) %>% round(digits = 1)`%). 

```{r}
# Compute a whole mess of distributions and their overlaps

############################
# Crowdsourced data, no mix
met.tmp <- met %>% filter(source == "cs" & scenario == "nomix_imbal")

# By tair
cs_nomix_tair_snow_dens <- with(met.tmp, density(tair[phase == "snow"], 
                           from = min(tair), 
                           to = max(tair)))
cs_nomix_tair_rain_dens <- with(met.tmp, density(tair[phase == "rain"], 
                           from = min(tair), 
                           to = max(tair)))
cs_nomix_tair_snowrain_joint <- pmin(cs_nomix_tair_snow_dens$y, cs_nomix_tair_rain_dens$y)
cs_nomix_tair_snowrain_overlap = (sum(cs_nomix_tair_snowrain_joint)/sum(cs_nomix_tair_snow_dens$y) +
                           sum(cs_nomix_tair_snowrain_joint)/sum(cs_nomix_tair_rain_dens$y))/2


# By twet
cs_nomix_twet_snow_dens <- with(met.tmp, density(twet[phase == "snow"], 
                           from = min(twet), 
                           to = max(twet)))
cs_nomix_twet_rain_dens <- with(met.tmp, density(twet[phase == "rain"], 
                           from = min(twet), 
                           to = max(twet)))
cs_nomix_twet_snowrain_joint <- pmin(cs_nomix_twet_snow_dens$y, cs_nomix_twet_rain_dens$y)
cs_nomix_twet_snowrain_overlap = (sum(cs_nomix_twet_snowrain_joint)/sum(cs_nomix_twet_snow_dens$y) +
                           sum(cs_nomix_twet_snowrain_joint)/sum(cs_nomix_twet_rain_dens$y))/2

############################
# Synoptic data, no mix
met.tmp <- met %>% filter(source == "nh" & scenario == "nomix_imbal")

# By tair
nh_nomix_tair_snow_dens <- with(met.tmp, density(tair[phase == "snow"], 
                           from = min(tair), 
                           to = max(tair)))
nh_nomix_tair_rain_dens <- with(met.tmp, density(tair[phase == "rain"], 
                           from = min(tair), 
                           to = max(tair)))
nh_nomix_tair_snowrain_joint <- pmin(nh_nomix_tair_snow_dens$y, nh_nomix_tair_rain_dens$y)
nh_nomix_tair_snowrain_overlap = (sum(nh_nomix_tair_snowrain_joint)/sum(nh_nomix_tair_snow_dens$y) +
                           sum(nh_nomix_tair_snowrain_joint)/sum(nh_nomix_tair_rain_dens$y))/2


# By twet
nh_nomix_twet_snow_dens <- with(met.tmp, density(twet[phase == "snow"], 
                           from = min(twet), 
                           to = max(twet)))
nh_nomix_twet_rain_dens <- with(met.tmp, density(twet[phase == "rain"], 
                           from = min(twet), 
                           to = max(twet)))
nh_nomix_twet_snowrain_joint <- pmin(nh_nomix_twet_snow_dens$y, nh_nomix_twet_rain_dens$y)
nh_nomix_twet_snowrain_overlap = (sum(nh_nomix_twet_snowrain_joint)/sum(nh_nomix_twet_snow_dens$y) +
                           sum(nh_nomix_twet_snowrain_joint)/sum(nh_nomix_twet_rain_dens$y))/2


############################
# Crowdsourced data, no mix, 0 to 4°C temp range
met.tmp <- met %>% filter(source == "cs" & scenario == "nomix_imbal" & tair >= 0 & tair <= 4)

# By tair
cs_nomix_tair_04_snow_dens <- with(met.tmp, density(tair[phase == "snow"], 
                           from = min(tair), 
                           to = max(tair)))
cs_nomix_tair_04_rain_dens <- with(met.tmp, density(tair[phase == "rain"], 
                           from = min(tair), 
                           to = max(tair)))
cs_nomix_tair_04_snowrain_joint <- pmin(cs_nomix_tair_04_snow_dens$y, cs_nomix_tair_04_rain_dens$y)
cs_nomix_tair_04_snowrain_overlap = (sum(cs_nomix_tair_04_snowrain_joint)/sum(cs_nomix_tair_04_snow_dens$y) +
                           sum(cs_nomix_tair_04_snowrain_joint)/sum(cs_nomix_tair_04_rain_dens$y))/2

############################
# Synoptic data, no mix, 0 to 4°C temp range
met.tmp <- met %>% filter(source == "nh" & scenario == "nomix_imbal" & tair >= 0 & tair <= 4)

# By tair
nh_nomix_tair_04_snow_dens <- with(met.tmp, density(tair[phase == "snow"], 
                           from = min(tair), 
                           to = max(tair)))
nh_nomix_tair_04_rain_dens <- with(met.tmp, density(tair[phase == "rain"], 
                           from = min(tair), 
                           to = max(tair)))
nh_nomix_tair_04_snowrain_joint <- pmin(nh_nomix_tair_04_snow_dens$y, nh_nomix_tair_04_rain_dens$y)
nh_nomix_tair_04_snowrain_overlap = (sum(nh_nomix_tair_04_snowrain_joint)/sum(nh_nomix_tair_04_snow_dens$y) +
                           sum(nh_nomix_tair_04_snowrain_joint)/sum(nh_nomix_tair_04_rain_dens$y))/2

```


```{r}
# Make a plot of the density overlaps
distros_combined <- bind_rows(
  data.frame(temp = rep(cs_nomix_tair_snow_dens$x, 3), 
             density = c(cs_nomix_tair_snow_dens$y, cs_nomix_tair_rain_dens$y, cs_nomix_tair_snowrain_joint),
             type = rep(c("All Snow", "All Rain", "Overlap"), each = length(cs_nomix_tair_snow_dens$x)),
             source = "cs",
             temp_type = "Air Temp."),
  data.frame(temp = rep(cs_nomix_twet_snow_dens$x, 3), 
             density = c(cs_nomix_twet_snow_dens$y, cs_nomix_twet_rain_dens$y, cs_nomix_twet_snowrain_joint),
             type = rep(c("All Snow", "All Rain", "Overlap"), each = length(cs_nomix_twet_snow_dens$x)),
             source = "cs",
             temp_type = "Wet Bulb"),
  data.frame(temp = rep(nh_nomix_tair_snow_dens$x, 3), 
             density = c(nh_nomix_tair_snow_dens$y, nh_nomix_tair_rain_dens$y, nh_nomix_tair_snowrain_joint),
             type = rep(c("All Snow", "All Rain", "Overlap"), each = length(nh_nomix_tair_snow_dens$x)),
             source = "nh",
             temp_type = "Air Temp."),
  data.frame(temp = rep(nh_nomix_twet_snow_dens$x, 3), 
             density = c(nh_nomix_twet_snow_dens$y, nh_nomix_twet_rain_dens$y, nh_nomix_twet_snowrain_joint),
             type = rep(c("All Snow", "All Rain", "Overlap"), each = length(nh_nomix_twet_snow_dens$x)),
             source = "nh",
             temp_type = "Wet Bulb")
) %>% 
  mutate(source_lab = ifelse(source == "cs",
                             "Crowdsourced", "Synoptic"))

density_overlap_plot<- 
  distros_combined %>% 
  filter(temp_type == "Air Temp.") %>% 
  ggplot(aes(temp, density, fill = type), alpha = 0.8) +
  geom_area(position = position_identity(), color = "black") +
  scale_fill_manual(values = hcl.colors(3, "Blue-Red2"),
                         name = "Phase") + 
  labs(x = "Air temperature (°C)", y = "Density") +
  facet_wrap(~source_lab, scales = "free")

density_overlap_plot

save_plot(filename = "figures/fig07_density_overlap.png", 
          plot = density_overlap_plot, 
          base_height = 4, 
          base_width = 9)

```

If we investigate the observed air temperature distributions of rain and snow, we find that the largest ranges in snowfall frequency predicted by the different methods (Figure 6) correspond roughly to the air temperatures with the greatest degree of overlap in rain and snow occurrence. Specifically, there is a `r (cs_nomix_tair_snowrain_overlap * 100) %>% round(digits = 1)`% overlap in the distributions of rain and snow as plotted by air temperature in the crowdsourced dataset and a `r (nh_nomix_tair_snowrain_overlap * 100) %>% round(digits = 1)`% overlap in the synoptic dataset. The maximum overlaps of the rain and snow distributions occur at `r distros_combined %>% filter(type == "Overlap" & source == "cs" & temp_type == "Air Temp.") %>% slice_max(density) %>% pull(temp) %>% round(digits = 1)`°C and `r distros_combined %>% filter(type == "Overlap" & source == "nh" & temp_type == "Air Temp.") %>% slice_max(density) %>% pull(temp) %>% round(digits = 1)`°C in the crowdsourced and synoptic datasets, respectively. What's more, if we only examine the 0°C-4°C range of poor performance, we find an astonishing `r (cs_nomix_tair_04_snowrain_overlap * 100) %>% round(digits = 1)`% overlap in the crowdsourced dataset and `r (nh_nomix_tair_04_snowrain_overlap * 100) %>% round(digits = 1)`% in the synoptic dataset. In other words, the air temperature characteristics of the two phases are nearly indistinguishable between 0°C and 4°C. 

```{r}

# Get the overlapped data and summarize by air temperature bin
overlap_dens_tair <- distros_combined %>% 
  filter(temp_type == "Air Temp." & type == "Overlap")
overlap_dens_tair_binned <- overlap_dens_tair %>% 
  mutate(tair_bin = round(temp/0.5) * 0.5) %>% 
  group_by(tair_bin, source_lab, source) %>% 
  summarise(density_av = mean(density)) 

# Get avg accuracy from all methods
accuracy_av_bytemp <- summary_bytemp %>% 
  filter(scenario == "nomix_imbal") %>% 
  group_by(tair_bin, source) %>% 
  summarize(accuracy_av_pct = mean(accuracy_pct),
            rain_bias_av_pct = mean(rain_bias_pct),
            snow_bias_av_pct = mean(snow_bias_pct))

# Join accuracy to overlap density
overlap_dens_tair_binned <- left_join(overlap_dens_tair_binned,
                                      accuracy_av_bytemp,
                                      by = c("tair_bin", "source"))

# Plot
overlap_accuracy_relationship_plot <- 
  plot_grid(
  plot_grid(
    ggplot(overlap_dens_tair_binned, aes(tair_bin, accuracy_av_pct, color = source_lab)) + 
      geom_line() +
      labs(y = "Average\naccuracy (%)") +
      theme(axis.title.x = element_blank(),
          axis.text.x = element_blank(),
          legend.position = "none") +
      scale_color_manual(name = "Source", values = hcl.colors(2, "Blue-Red2")),
    ggplot(overlap_dens_tair_binned, aes(tair_bin, density_av, color = source_lab)) +
      geom_line() +
      labs(x = "Air temperature (°C)",
          y = "Overlap\ndensity") +
      theme(legend.position = "none")+ 
      scale_color_manual(name = "Source", values = hcl.colors(2, "Blue-Red2")),
  ncol = 1, align = "vh", labels = "auto"
  ),
  ggplot(overlap_dens_tair_binned, aes(density_av, accuracy_av_pct, 
                                       color = source_lab, shape = source_lab)) + 
    geom_smooth(method = "lm", se = F, lty = "dashed", alpha = 0.7) +
    geom_point() +
    theme(legend.position = "top") +
    labs(x = "Overlap density",
         y = "Average Accuracy (%)") +
    scale_color_manual(values = hcl.colors(2, "Blue-Red2"))+
    guides(color = guide_legend(title.position = "top", title = "Dataset",
                                legend.justification = "center", legend.title.align = 0.5),
           shape = guide_legend(title.position = "top", title = "Dataset",
                                legend.justification = "center", legend.title.align = 0.5)),
  ncol = 2, labels = c("a", "c")
)

overlap_accuracy_relationship_plot

save_plot(filename = "figures/fig08_overlap_accuracy_relationship.png", 
          plot = overlap_accuracy_relationship_plot, 
          base_height = 5, 
          base_width = 9)

```
```{r}
# Quantify the relationship
cs_overlap_accuracy_lm <- 
  overlap_dens_tair_binned %>% 
  filter(source == "cs") %>% 
  lm(accuracy_av_pct ~ density_av, data = .) %>% 
  summary()

# Calculate the relationship
nh_overlap_accuracy_lm <- 
  overlap_dens_tair_binned %>% 
  filter(source == "nh") %>% 
  lm(accuracy_av_pct ~ density_av, data = .) %>% 
  summary()



```
This lack of differentiability at temperatures from 0° to 4°C reveals an issue with using near-surface meteorology to partition rain and snow. The left panels of Figure 8 below present the average accuracy of the benchmark and machine learning phase partitioning methods plus the overlap in rain and snow distributions from the two datasets. We plot both variables against air temperature to demonstrate how the dip in performance corresponds to the increase in overlap. The right panel then shows that there is a statistically significant (p < 0.0005) negative relationship between the distribution overlap and average accuracy, where accuracy decreases as overlap increases. Using ordinary least squares regression, we compute an r^2^ of `r cs_overlap_accuracy_lm$r.squared %>% round(digits = 2)` for this relationship in the crowdsourced dataset and `r nh_overlap_accuracy_lm$r.squared %>% round(digits = 2)` in the synoptic. That means that the air temperature distribution overlap between rain and snow explains at least `r min(c(cs_overlap_accuracy_lm$r.squared, nh_overlap_accuracy_lm$r.squared))*100 %>% round(digits = 0)` % of the variance in precipitation phase partitioning accuracy. 

This relationship (Figure 6) provides information on the limits of using near-surface meteorology to partition rain and snow. Recall that the benchmark and machine learning methods, when applied to the synoptic dataset, expressed higher accuracy values than in the crowdsourced dataset (Table 1). Consistent with the findings in this section, the higher accuracy in the synoptic dataset is complemented by a lower percentage of rain and snow distribution overlap (`r (nh_nomix_tair_snowrain_overlap * 100) %>% round(digits = 1)`%) than in the crowdsourced dataset (`r (cs_nomix_tair_snowrain_overlap * 100) %>% round(digits = 1)`%). This, plus the fact that the air temperature distribution overlap between rain and snow explains at least 73% of the variance in precipitation phase partitioning accuracy, indicates that the key limiting factor to performance is the overlap of the air temperature distributions of the different phases. The higher the overlap, the worse the performance. Even the machine learning methods are beholden to this phenomenon (Figure 2, Table 2). Thus, as of now, regions with a high degree of rain-snow air temperature distribution overlap (i.e., those with characteristics of the crowdsourced dataset) are likely to see minimum benchmark accuracy values approaching `r summary_bytemp %>% filter(scenario == "nomix_imbal" & source == "cs" & ppm == "thresh_twet_1") %>% slice_min(accuracy_pct) %>% pull(accuracy_pct) %>% round(digits = 1)`%, while regions with less overlap (i.e., those with characteristics of the synoptic dataset) may see improved minimum accuracies, approaching `r summary_bytemp %>% filter(scenario == "nomix_imbal" & source == "nh" & ppm == "thresh_twet_0.5") %>% slice_min(accuracy_pct) %>% pull(accuracy_pct) %>% round(digits = 1)`%. Although the machine learning methods may moderately improve this performance dip in the crowdsourced (up to `r filter(summary_bytemp, source == "cs" & scenario == "nomix_imbal" & ppm == "nn") %>% pull(accuracy_pct) %>% min() %>% round(digits = 1)`%) and synoptic datasets (up to `r filter(summary_bytemp, source == "nh" & scenario == "nomix_imbal" & ppm == "rf") %>% pull(accuracy_pct) %>% min() %>% round(digits = 1)`%) , they do not eliminate it, demonstrating a limit to precipitation phase partitioning accuracy using near-surface meteorology.

# Discussion

## Variable importance
As noted in the results section covering benchmark performance, the inclusion of humidity, typically in the form of wet bulb or dew point temperature, generally increases accuracy compared to methods that only utilize air temperature. This is underscored by the higher performance of the machine learning methods, both of which include wet bulb temperature, dew point temperature, and relative humidity as predictor variables. By using two tree-based machine learning methods, we are able to evaluate variable importance, which represents how much including a given variable increases performance, or, conversely, how much removing it decreases performance. Figure 9 shows the variable importance plots for the XGBoost and random forest models trained on the crowdsourced and synoptic datasets. Again we see the primacy of wet bulb temperature as an accurate way of predicting precipitation phase—in every plot it is the most important variable. In the crowdsourced dataset, dew point temperature is the second most important variable for both machine learning methods, while air temperature ranks second in the synoptic dataset. Relative humidity was the least important variable in three of the four plots.


```{r}
vip_plot <-
  plot_grid(
  plot_grid(
    ggdraw() + draw_label(""),
    ggdraw()+draw_label("XGBoost"),
    ggdraw()+draw_label("Random forest"),
    nrow = 1, rel_widths = c(0.1, 1, 1)
  ),
  plot_grid(
    ggdraw()+draw_label("Crowdsourced", angle = 90),
    ml_for_vip[[1]] %>% vip::vip(include_type = T, 
                              mapping = aes(fill = .data[["Importance"]]),
                              aesthetics = list(color = "grey35", size = 0.2)) + 
      scale_x_discrete(labels = c(expression(T[w]), expression(T[d]), 
                              expression(T[a]), "RH"),
                   breaks = c("twet", "tdew", "tair", "rh")) +
      scale_fill_gradientn(colors = hcl.colors(3, "Heat2"),
                       labels = c("Higher", "Lower"),
                       breaks = c(0.9,0.1)) +
      theme(legend.position = c(0.7, 0.4)),
    ml_for_vip[[2]]%>%  vip::vip(include_type = T, 
                                 mapping = aes(fill = .data[["Importance"]]),
                                 aesthetics = list(color = "grey35", size = 0.2))+ 
      scale_x_discrete(labels = c(expression(T[w]), expression(T[d]), 
                                  expression(T[a]), "RH"),
                       breaks = c("twet", "tdew", "tair", "rh")) +
      scale_fill_gradientn(colors = hcl.colors(3, "Heat2")),
    nrow = 1, rel_widths = c(0.1, 1, 1), labels = c("", "a", "b")
  ),
  plot_grid(
    ggdraw()+draw_label("Synoptic", angle = 90),
    ml_for_vip[[3]]%>%  vip::vip(include_type = T, 
                                 mapping = aes(fill = .data[["Importance"]]),
                                 aesthetics = list(color = "grey35", size = 0.2))+ 
      scale_x_discrete(labels = c(expression(T[w]), expression(T[d]), 
                                  expression(T[a]), "RH", "P"),
                       breaks = c("twet", "Dewpoint", "Air_Temp", "RH",
                                  "gridded_data_pres")) +
      scale_fill_gradientn(colors = hcl.colors(3, "Heat2")),
    ml_for_vip[[4]]%>%  vip::vip(include_type = T, 
                                 mapping = aes(fill = .data[["Importance"]]),
                                 aesthetics = list(color = "grey35", size = 0.2))+ 
      scale_x_discrete(labels = c(expression(T[w]), expression(T[d]), 
                                  expression(T[a]), "RH", "P"),
                       breaks = c("twet", "Dewpoint", "Air_Temp", "RH",
                                  "gridded_data_pres")) +
      scale_fill_gradientn(colors = hcl.colors(3, "Heat2")),
    nrow = 1, rel_widths = c(0.1, 1, 1), labels = c("", "c", "d")
  ),
  nrow = 3, rel_heights = c(0.2, 1, 1))

vip_plot

save_plot(filename = "figures/fig09_vip.png", 
          plot = vip_plot, 
          base_height = 7, 
          base_width = 9)
  
```

## Class imbalances

```{r}
phase_by_scenario <- met %>%  
  group_by(source, scenario, phase) %>% 
  summarize(n = n()) %>% 
  mutate(freq = n/sum(n) * 100)
```

We had more snow (`r phase_by_scenario %>% filter(source == "cs" & scenario == "nomix_imbal" & phase == "snow") %>% pull(freq) %>% round(digits = 1)`%) than rain (`r phase_by_scenario %>% filter(source == "cs" & scenario == "nomix_imbal" & phase == "rain") %>% pull(freq) %>% round(digits = 1)`%) in the testing split of the crowdsourced data without mixed precipitation. When we included mixed precipitation, the breakdown was `r phase_by_scenario %>% filter(source == "cs" & scenario == "allphase_imbal" & phase == "snow") %>% pull(freq) %>% round(digits = 1)`% snow, `r phase_by_scenario %>% filter(source == "cs" & scenario == "allphase_imbal" & phase == "rain") %>% pull(freq) %>% round(digits = 1)`% rain, and `r phase_by_scenario %>% filter(source == "cs" & scenario == "allphase_imbal" & phase == "mix") %>% pull(freq) %>% round(digits = 1)`% mixed. Our data preparation workflow did not use any over- or undersampling method to adjust class imbalances, so we reran the random forest and XGBoost hyperparameter tuning exercise (supplemental material) for the no mixed and all phase scenarios. We found that using the synthetic minority oversampling technique (SMOTE) (Chawla et al., 2002) that accuracy declined in every case relative to using the imbalanced data directly.

```{r}
summary_all %>% 
  filter(source == "cs" & type == "ml") %>% 
  mutate(scenario2 = ifelse(str_detect(scenario, "imbal"), 
                             "Imbalanced", "SMOTE"),
         phases = ifelse(str_detect(scenario, "nomix"),
                         "No Mixed", "All Phases"),
         across(where(is.numeric), round, digits = 1)) %>% 
  ungroup() %>% 
  select(PPM = ppm, 
         `Class balance` = scenario2, 
         Phases = phases, 
         `Accuracy (%)` = accuracy_pct, 
         `Rain bias (%)` = rain_bias_pct,
         `Snow bias (%)` = snow_bias_pct,
         `Mixed bias (%)` = mixed_bias_pct) %>% 
  knitr::kable()
```

Although we did not find that oversampling the data to correct class imbalances improved the machine learning predictions, such imbalances may partly explain performance differences between the two datasets. The synoptic dataset had a more evenly balanced set of snow (`r phase_by_scenario %>% filter(source == "nh" & scenario == "nomix_imbal" & phase == "snow") %>% pull(freq) %>% round(digits = 1)`%) than rain (`r phase_by_scenario %>% filter(source == "nh" & scenario == "nomix_imbal" & phase == "rain") %>% pull(freq) %>% round(digits = 1)`%) observations than the crowdsourced dataset. Another potential explanation is that there are three orders of magnitude more observations in the synoptic dataset than the crowdsourced dataset, meaning there are more data that the machine learning algorithms can use to determine relationships between near-surface meteorology and precipitation phase. The most likely explanation, however, is the difference in rain-snow air temperature distribution overlaps between the two datasets (Figure 7). Mountain Rain or Snow specifically targets mountain regions and storm events where accurate phase predictions are most difficult to produce, giving the crowdsourced rain and snow data a `r (cs_nomix_tair_snowrain_overlap * 100) %>% round(digits = 1)`% overlap by air temperature and the synoptic data a `r (nh_nomix_tair_snowrain_overlap * 100) %>% round(digits = 1)`% overlap. Given the strong relationship between overlap and accuracy (Figure 8), the lower overlap in the synoptic dataset likely explains the higher performance of the machine learning and benchmark methods.

