---
title: "Mountain Rain or Snow 2020-2021 Analysis"
author: "Keith Jennings"
date: "7/26/2021"
output: github_document
---

# Introduction

The `Analysis` directory includes the processing scripts used to evaluate precipitation phase data submitted by citizen scientists taking part in _Mountain Rain or Snow_. This includes pre-processing the observations, downloading ancillary meteorological data, estimating air temperature at each observation point, and quality controlling the precipitation phase reports.

The analysis below includes observations submitted near Lake Tahoe during the 2020 and 2021 campaigns.

# Step 1: Data Pre-Processing

The `mros_cit_sci_obs_preprocess.R` script in `analysis/pre_processing` associates the following geospatial data with each observation:

* Level II, III, and IV Ecoregions
* US State
* Elevation (m)

The script also formats the time data and removes all observations submitted without geolocation info.

# Step 2: Meteorological Data Download and Pre-Processing

Few of the citizen science observations are submitted near existing meteorological measurement stations, meaning we need to model the air temperature for each rain, snow, mixed data point. We do this with air temperature data from four networks:

Network Name | URL
-------------|---------------------------------------------------------
HADS         | <https://mesonet.agron.iastate.edu/request/dcp/fe.phtml>
SNOTEL       | <https://wcc.sc.egov.usda.gov/reportGenerator/>
RAWS         | <https://raws.dri.edu/>
NCDC LCD     | <https://gis.ncdc.noaa.gov/maps/ncei/lcd>

The two intial datasets can be accessed programatically using `analysis/download/data_download_hads.R` and `analysis/download/data_download_snotel.R` while the latter two need to be downloaded manually from the links. After downloading the data, we prepare them for the temperature model using `analysis/pre_processing/mros_met_data_preprocess.R`.

# Step 3: Air Temperature Modeling

We then model the air temperature at each location using four methods in the `analysis/met_modeling/mros_air_temperature_model.R` script:

* Inverse distance weighting (IDW), plus a constant lapse rate of -0.005°C/m
* IDW, plus a variable lapse rate computed from air temperature observations
* Nearest met station, plus a constant lapse rate of -0.005°C/m
* Nearest met station, plus a variable lapse rate computed from air temperature observations

As a result of its higher performance in terms of mean bias and r<sup>2</sup>, we use the IDW plus variable lapse rate method when assigning air temperature to each observation point.

Future versions of the code will model additional meteorological variables, such as relative humidity, dew point temperature, and wet bulb temperature.

# Step 4: Quality Control of Citizen Science Observations

Next, we flag observations based on five criteria:

1. Whether precipitation was recorded that day at the nearby NCDC LCD met stations.
2. The estimated air temperature relative to realistic values for rain and snow.
3. The relative humidity at the observation point
4. Average distance from the met stations.
5. If the timestamp was a duplicate of an observation from the same observer.

We then output the processed, quality controlled observations to a shareable data file: `data/processed/mros_obs_processed_2020_2021.RDS`

# Looking at the data

Now we can examine the data more in depth. First, we'll start an R session and import the file:

```{r message = FALSE}
# Load the tidyverse
library(tidyverse)

# Use cowplot for plot formatting and import plot styles
library(cowplot); theme_set(theme_cowplot())
source("functions/mros_plot_formats.R")

# Lubridate for date handling tools
library(lubridate)

# Import data
obs <- readRDS("../data/processed/mros_obs_processed_2020_2021.RDS")

# Add factor level to phase to force Rain > Mixed > Snow order
obs <- obs %>% mutate(phase = factor(phase, levels = c("Rain", "Mixed", "Snow")))
```

Within the `obs` file we have a total of `r length(obs$phase)` observations submitted between `r min(obs$date)` and `r max(obs$date)` in the Lake Tahoe area. Of these `r length(obs$phase)` observations, `r length(filter(obs, tair_flag == "Pass" & ppt_flag == "Pass" & rh_flag == "Pass" & dist_flag == "Pass" & dupe_flag == "Pass")$phase)` passed all of the quality control checks (this is `r round(length(filter(obs, tair_flag == "Pass" & ppt_flag == "Pass" & rh_flag == "Pass" & dist_flag == "Pass" & dupe_flag == "Pass")$phase)/length(obs$phase) * 100, digits = 1)`% of the database).

For the rest of document, we'll evaluate only data that passed the QC checks. We'll also add elevation bins for additional analyses

```{r message = FALSE}
# Filter to passing obs
obsPass <- filter(obs, tair_flag == "Pass" & 
                        ppt_flag == "Pass" & 
                        rh_flag == "Pass" &
                        dist_flag == "Pass" & 
                        dupe_flag == "Pass")

# Add elevation info
obsPass <- obsPass %>% 
   mutate(elev_bin = cut_width(elev, width = 500))
```

## Observations by elevation

We received precipitation reports in the study area between `r min(obsPass$elev)` m and `r max(obsPass$elev)` m, with mean and median report elevations of `r round(mean(obsPass$elev), digits = 0)` m and `r round(median(obsPass$elev), digits = 0)` m, respectively. In total, there were `r length(filter(obsPass, phase == "Snow")$phase)`, `r length(filter(obsPass, phase == "Rain")$phase)`, `r length(filter(obsPass, phase == "Mixed")$phase)`, snow, rain, and mixed observations. 

The elevational breakdown of precipitation phase looks like this:

```{r echo = FALSE}
ggplot(obsPass, aes(elev_bin, fill = phase)) +
  geom_bar(position = "fill", color = "black") +
  phase_fill_scale +
  labs(x = "Elevation Bin (m)",
       y = "Type Proportion (%)") +
  scale_y_continuous(breaks = seq(0, 1, by = 0.25),
                     labels = c("0", "25", "50", "75", "100")) +
  scale_x_discrete(labels = c("< 1250",
                              "1250 - 1750",
                              "1750 - 2250",
                              "> 2250"))
```

Similarly, we can look at the number of reports per precipitation phase by elevation:

```{r echo = FALSE, warning = FALSE, message = FALSE, fig.dim= c(6,6)}
ggplot(obsPass, aes(elev, fill = phase)) + 
  geom_histogram(color = "black") + 
  facet_wrap(~phase) + 
  coord_flip() + 
  phase_fill_scale +
  labs(y = "Number of Observations", 
       x = "Elevation (m)") +
  theme(strip.background = element_rect(fill = "white", color = "black"),
        legend.position = "none")
```

Or grouped together:

```{r echo = FALSE, warning = FALSE, message = FALSE, fig.dim= c(6,6)}
ggplot(obsPass, aes(elev, fill = phase)) + 
  geom_histogram(color = "black") + 
  coord_flip() + 
  phase_fill_scale +
  labs(y = "Number of Observations", 
       x = "Elevation (m)") +
  theme(strip.background = element_rect(fill = "white", color = "black"),
        legend.position = c(0.8, 0.8))
```

## Observations by state and ecoregion

Across the study period, we had the following geographic breakdown of observations.

By state:

```{r echo = FALSE, warning = FALSE, message = FALSE}
obsPass %>% 
  group_by(state) %>% 
  summarise(n = n()) %>% 
  knitr::kable()
```

By US EPA Level III Ecoregion:

```{r echo = FALSE, warning = FALSE, message = FALSE}
obsPass %>% 
  group_by(eco_l3) %>% 
  summarise(n = n()) %>% 
  knitr::kable()
```

And by US EPA Level IV Ecoregion:

```{r echo = FALSE, warning = FALSE, message = FALSE}
obsPass %>% 
  group_by(eco_l4) %>% 
  summarise(n = n()) %>% 
  knitr::kable()
```

## Observations by year

Our *Tahoe Rain or Snow* study period comprised two water years: 2020 and 2021.

```{r message = FALSE}
# Add water year info
obsPass <- obsPass %>% 
   mutate(wy = ifelse(month(date) >= 10,
                      year(date) + 1,
                      year(date)),
          dowy = ifelse(month(date) >= 10,
                        yday(date) - 273,
                        yday(date) + 92))

# Compute cumulative observations per phase and water year
obsPass <- obsPass %>% 
  arrange(date) %>% 
  group_by(wy, phase) %>% 
  mutate(n_cumulative = row_number())
```

We received `r length(filter(obsPass, wy == 2020)$phase)` observations in water year 2020 and `r length(filter(obsPass, wy == 2021)$phase)` in 2021. Snow was the most frequent phase in both water years, but both rain and mixed precipitation increased in relative proportion in 2021. 

```{r echo = FALSE, warning = FALSE, message = FALSE}
obsPass %>% 
  group_by(wy, phase) %>% 
  summarise(n = n()) %>% 
  ungroup() %>% 
  group_by(wy) %>% 
  mutate(pct = round(n / sum(n) * 100, digits = 1)) %>% 
  knitr::kable()
```


```{r echo = FALSE, warning = FALSE, message = FALSE, fig.dim= c(6,4)}
ggplot(obsPass, aes(dowy, n_cumulative, color = phase)) + 
  geom_line(lwd = 1) + 
  labs(x = "Date",
       y = "Cumulative Observations") +
  theme(legend.position = c(0.1, 0.9), legend.title = element_blank(),
        strip.background = element_rect(fill = NA, color = "black")) +
  scale_color_manual(values = c("blue", "purple", "black")) +
  facet_wrap(~wy, ncol = 1) +
  scale_x_continuous(breaks = c(1, 93, 183, 274), 
                     labels = c("Oct. 1", "Jan. 1", "Apr. 1", "Jul. 1"))
```
## Observations by air temperature

Modeled meteorological data show volunteers submitted precipitation phase reports from a minimum air temperature value of `r round(min(obsPass$tair), digits = 1)`°C to a maximum of `r round(max(obsPass$tair), digits = 1)`°C with a median value of `r round(median(obsPass$tair), digits = 1)`°C. What’s more, the vast majority of reports came from a relatively narrow air temperature range. Our data show 95% of observations corresponded to air temperatures between `r round(quantile(obsPass$tair, probs = c(0.025, 0.975))[1], digits = 1)`°C and `r round(quantile(obsPass$tair, probs = c(0.025, 0.975))[2], digits = 1)`°C. 

```{r echo = FALSE, warning = FALSE, message = FALSE}
plot_grid(
  ggplot(obsPass, aes(tair, color = phase)) +
    geom_density(lwd = 1)+
    scale_color_manual(values = c("blue", "purple", "black")) +
    labs(x = "Air Temperature (°C)", y = "Density") +
    theme(legend.position = c(0.8,0.8), legend.title = element_blank()),
  ggplot(obsPass, aes(twet, color = phase)) +
    geom_density(lwd = 1)+
    scale_color_manual(values = c("blue", "purple", "black")) +
    labs(x = "Wet Bulb Temperature (°C)", y = "Density") +
    theme(legend.position = "none"),
  cols = 1, align = "vh", labels = "auto"
)
```
## Rain-snow partitioning

Using the observations and modeled meteorological data, we can compute snowfall probability curves and 50% snowfall probability temperatures, the latter of which are used as thresholds in models to split solid and liquid precipitation. We provide a methods overview here and the full code can be found in `mros_obs_analysis.R`. In short, we bin air temperature in 1°C increments from -10°C to 20°C and wet bulb temperature in 1°C increments from -12°C to 16°C. We then compute the probability of snowfall occurring in each air and wet bulb temperature bin. Those data are next fit with a hyperbolic tangent as in Dai (2008) and Jennings et al. (2018) to create snowfall probability curves. The 50% snowfall probability air and wet bulb temperature thresholds are where the fitted curves pass the 50% mark.

Let's import the data.

```{r message = FALSE}

# Import data
rs_p <- readRDS("../data/processed/mros_obs_rs_partitioning_2020_2021.RDS")

```

We can then look at the air temperature snowfall probability plot:

```{r echo = FALSE, warning = FALSE, message = FALSE, fig.dim= c(6,4)}
ggplot(rs_p[["snow_pred_tair"]], aes(tair_bin_num, snow_prob_pred * 100)) + 
  geom_line(lwd = 1) + 
  geom_hline(yintercept = 50, color = "gray", lty = "dashed", lwd = 1) +
  geom_line(data = rs_p[["snow_prob_tair"]], aes(tair_bin_num, snow_prob * 100), 
             color = "darkgreen", lwd = 1) +
  labs(x = expression("Air Temperature ("*degree*C*")"),
       y = "Snowfall Probability (%)") +
  annotate(geom = "text", x = -8, y = 25, label = "Fitted curve", hjust = 0) +
  annotate(geom = "text", x = -8, y = 15, label = "Probability by bin", color = "darkgreen", hjust = 0) +
  annotate(geom = "text", x = 10, y = 50, label = paste0("50% probability = ", round(rs_p[["temp50_tair"]], digits = 1), "°C"), 
           color = "gray27", hjust = 0, vjust = -.5)
```

The wet bulb temperature snowfall probability plot:

```{r echo = FALSE, warning = FALSE, message = FALSE, fig.dim= c(6,4)}
ggplot(rs_p[["snow_pred_twet"]], aes(twet_bin_num, snow_prob_pred * 100)) + 
  geom_line(lwd = 1) + 
  geom_hline(yintercept = 50, color = "gray", lty = "dashed", lwd = 1) +
  geom_line(data = rs_p[["snow_prob_twet"]], aes(twet_bin_num, snow_prob * 100), 
             color = "aquamarine3", lwd = 1) +
  labs(x = expression("Wet Bulb Temperature ("*degree*C*")"),
       y = "Snowfall Probability (%)") +
  annotate(geom = "text", x = -10, y = 25, label = "Fitted curve", hjust = 0) +
  annotate(geom = "text", x = -10, y = 15, label = "Probability by bin", color = "aquamarine3", hjust = 0) +
  annotate(geom = "text", x = 6, y = 50, label = paste0("50% probability = ", round(rs_p[["temp50_twet"]], digits = 1), "°C"), 
           color = "gray27", hjust = 0, vjust = -.5)
```

And the two of them combined:

```{r echo = FALSE, warning = FALSE, message = FALSE, fig.dim= c(6,4)}
ggplot() + 
  geom_hline(yintercept = 50, color = "gray", lty = "dashed", lwd = 1) +
  geom_line(data = rs_p[["snow_prob_tair"]], aes(tair_bin_num, snow_prob * 100), 
             color = "darkgreen", lwd = 1) +
  geom_line(data = rs_p[["snow_prob_twet"]], aes(twet_bin_num, snow_prob * 100), 
             color = "aquamarine3", lwd = 1) +
  labs(x = expression("Temperature ("*degree*C*")"),
       y = "Snowfall Probability (%)") +
  annotate(geom = "text", x = -10, y = 25, label = "Air temperature", color = "darkgreen", hjust = 0) +
  annotate(geom = "text", x = -10, y = 15, label = "Wet bulb temperature", color = "aquamarine3", hjust = 0)
```

We notice the 50% threshold is a fair bit higher for air temperature verus wet bulb temperature, which is expected because the former is always warmer than the latter when relative humidity is less than 100%. We also see a fair bit more noise in the curve for air temperature between approximately 25% and 50%. 

```{r echo = FALSE, warning = FALSE, message = FALSE}
tmp <- filter(obsPass, tair >= min(filter(rs_p[["snow_prob_tair"]],snow_prob <= .50 & snow_prob >= .25)$tair_bin_num) &
                tair <= max(filter(rs_p[["snow_prob_tair"]],snow_prob <= .50 & snow_prob >= .25)$tair_bin_num))
rs_hum <- tmp %>% 
  group_by(phase) %>% 
  summarise("mean_rh" = mean(rh)) %>% 
  mutate(temp = "air")
tmp <- filter(obsPass, twet >= min(filter(rs_p[["snow_prob_twet"]],snow_prob <= .50 & snow_prob >= .25)$twet_bin_num, na.rm = T) &
                twet <= max(filter(rs_p[["snow_prob_twet"]],snow_prob <= .50 & snow_prob >= .25)$twet_bin_num, na.rm = T))

rs_hum <- bind_rows(
  rs_hum,
  tmp %>% 
  group_by(phase) %>% 
  summarise("mean_rh" = mean(rh)) %>% 
  mutate(temp = "wet bulb"))

knitr::kable(rs_hum)

```

## 

```{r echo = FALSE, warning = FALSE, message = FALSE, fig.dim= c(6,3)}
p <- ggplot(filter(obsPass, phase == "Snow"), aes(tair)) + geom_density(fill = "gray") +
  labs(x = "Air Temperature (°C)", y = "Density")
p
# Get the density data from the plot
# This will allow us to look at right/wrong values
d <- ggplot_build(p)$data[[1]]
```

```{r echo = FALSE, warning = FALSE, message = FALSE, fig.dim= c(6,3)}
tair_thresh = 0
pct_incorrect = length(filter(obsPass, tair>tair_thresh & phase == "Snow")$phase) /length(filter(obsPass, phase == "Snow")$phase) * 100
p +  geom_area(data = filter(d, x>tair_thresh), aes(x=x,y=y), fill = "darkred", alpha = 0.5) +
  annotate(geom = "text", x = 2.5, y = 0.1, 
           label = paste0(round(pct_incorrect, digits = 1), "% of snow misidentified"), color = "darkred",
           hjust = 0)
```

```{r echo = FALSE, warning = FALSE, message = FALSE, fig.dim= c(6,3)}
tair_thresh = 2.2
pct_incorrect = length(filter(obsPass, tair>tair_thresh & phase == "Snow")$phase) /length(filter(obsPass, phase == "Snow")$phase) * 100
p +  geom_area(data = filter(d, x>tair_thresh), aes(x=x,y=y), fill = "darkred", alpha = 0.5) +
  annotate(geom = "text", x = 2.5, y = 0.1, 
           label = paste0(round(pct_incorrect, digits = 1), "% of snow misidentified"), color = "darkred",
           hjust = 0)
```

```{r echo = FALSE, warning = FALSE, message = FALSE, fig.dim= c(6,3)}
tair_thresh = 3.9
pct_incorrect = length(filter(obsPass, tair>tair_thresh & phase == "Snow")$phase) /length(filter(obsPass, phase == "Snow")$phase) * 100
p +  geom_area(data = filter(d, x>tair_thresh), aes(x=x,y=y), fill = "darkred", alpha = 0.5) +
  annotate(geom = "text", x = 2.5, y = 0.1, 
           label = paste0(round(pct_incorrect, digits = 1), "% of snow misidentified"), color = "darkred",
           hjust = 0)
```

## GPM IMERG comparison

```{r message = FALSE}

# Import data
gpm <- readRDS("../data/processed/mros_gpm_processed_2020_2021.RDS")

# Join the data
obsGPM <- left_join(obsPass,
                    gpm,
                    by = "id")

###############################################################################
# Summarize correct observations by temp bin

# Make wider bins for analysis
tair_max = 14
tair_min = -7
tair_bin_width = 1
obsGPM$tair_bin2 <- cut(obsGPM$tair, 
                     breaks = seq(tair_min, 
                                  tair_max, 
                                  by = tair_bin_width))


# Add numeric version of each bin
tair_cuts_to_number2 <- data.frame("tair_bin2" = levels(obsGPM$tair_bin2),
                                   "tair_bin_num2" = seq(tair_min + (0.5 * tair_bin_width),
                                                         tair_max - (0.5 * tair_bin_width), 
                                                         by = tair_bin_width))

# Join
obsGPM <- left_join(obsGPM, tair_cuts_to_number2, by = "tair_bin2")


# Add GPM probability thresholds for rain, snow, mixed
prob_thresh_upper_rain = 100
prob_thresh_lower_rain = 50
prob_thresh_upper_snow = 50
prob_thresh_lower_snow = 0
prob_thresh_upper_mixed = prob_thresh_lower_rain
prob_thresh_lower_mixed = prob_thresh_upper_snow

# Denote whether phase designation was correct or not
obsGPM$phase2 <- as.character(obsGPM$phase)
obsGPM <- obsGPM %>% 
  mutate(phase_noMIXED = case_when(phase2 == "Mixed" ~ "Rain",
                                   TRUE ~ phase2),
         gpm_phase = case_when(gpm_prob <= prob_thresh_upper_snow &
                                 gpm_prob >= prob_thresh_lower_snow ~ "Snow",
                               gpm_prob <= prob_thresh_upper_rain &
                                 gpm_prob >= prob_thresh_lower_rain ~ "Rain",
                               gpm_prob < prob_thresh_upper_mixed &
                                 gpm_prob > prob_thresh_lower_mixed~ "Mixed"),
         gpm_score = case_when(gpm_phase == phase2 ~ 1,
                               gpm_phase != phase2 ~ 0),
         gpm_score_noMIXED = case_when(gpm_phase == phase_noMIXED ~ 1,
                                       gpm_phase != phase_noMIXED ~ 0))

# Summarize by tair bin
gpm_summary_noMIXED <- obsGPM %>% 
  filter(!is.na(gpm_score_noMIXED)) %>% 
  group_by(tair_bin_num2) %>% 
  summarize(n_obs = length(gpm_score_noMIXED), 
            gpm_perf_pct = (sum(gpm_score_noMIXED) / n_obs) * 100,
            snow_pct_obs = (sum(phase == "Snow") / n_obs) * 100,
            snow_pct_gpm = (sum(gpm_phase == "Snow") / n_obs) * 100)

```

Then plot the analyzed results:

```{r echo = FALSE, warning = FALSE, message = FALSE, fig.dim= c(7,4.5)}
reshape2::melt(select(gpm_summary_noMIXED, -n_obs), id.vars = "tair_bin_num2", 
               variable.name = "metric", value.name = "pct") %>% 
  ggplot(., aes(tair_bin_num2, pct, color = metric)) +
  geom_line(lwd = 1) +
  labs(x = expression("Air Temperature Bin ("*degree*C*")"),
       y = "GPM Performance and Snow Probability (%)") + 
  scale_color_manual(values = c("black", "darkgreen", "blue"),
                     labels = c("GPM Success Rate", "Obs. Snow Prob.", 
                                "GPM Snow Prob."),
                     name = "") +
  theme(legend.position = c(0.1, 0.2))
```

## Rain-snow line comparison

Import the data

```{r message = FALSE}

library(lubridate) # for datetime handling

# Import data
gpm <- readRDS("../data/processed/gpm_rain_snow_line_2020_2021.RDS") %>% 
  mutate(datetime = with_tz(datetime, "Etc/Gmt+8")) # convert to PST
flr <- read.csv("../data/processed/CFF_2019_2021.csv", na.strings = " NaN") %>% 
  mutate(datetime = with_tz(as.POSIXct(paste0(year, "-", month, "-", day, " ", hour),
                               format = "%Y-%m-%d %H",
                               tz = "GMT"),
                            "Etc/Gmt+8"), # convert to PST
         rain_snow_line = BBH * 1000) # convert from km to m
obs <- readRDS("../data/processed/mros_obs_rain_snow_line_2020-2021.RDS") %>% 
  mutate(datetime_min = as.POSIXct(paste0(date, " 08:00"), 
                                   tz = "Etc/Gmt+8"),
         datetime_max = as.POSIXct(paste0(date, " 20:00"), 
                                   tz = "Etc/Gmt+8"))

# Join data
all <- left_join(gpm, flr, by = "datetime")

```

Plot an event:

```{r echo = FALSE, warning = FALSE, message = FALSE, fig.dim= c(7,4.5)}
# Plot an event
min_datetime = as.POSIXct("2020-04-04 05:00", tz = "Etc/Gmt+8")
max_datetime = as.POSIXct("2020-04-07 00:00", tz = "Etc/Gmt+8")
filter(all, datetime >= min_datetime & 
         datetime <= max_datetime) %>% 
  ggplot() + 
  geom_line(aes(datetime, rs_line)) + 
  geom_point(aes(datetime, rain_snow_line), color = "red") + 
  ylim(900, 3000) + 
  geom_segment(data = filter(obs, datetime_min >= min_datetime &
                               datetime_max <= max_datetime), 
               aes(x = datetime_min, y = rs_line, xend = datetime_max, yend = rs_line), 
               color = "purple", lwd = 1) + 
  labs(x = "Datetime",
       y = "Rain-Snow Line Elevation (m)")
```
